---
title: |
  | Financial Econometric in R/Python 
  | Assignment One
author: |
  | Group 2
  | Hessa Alabbas 02513615
  | Xin Zhan 02299544
  | Alex Rached 01894052
  | Yan Cai 02381303
  | Kexin Liu 02362049
  |
  | The Business School, Imperial College London
  | 
date: "23-10-2023"
output:
  pdf_document: 
    keep_tex: true›
    latex_engine: xelatex

header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{caption}
---


\newpage
\tableofcontents
\newpage

\pagebreak

# Introduction

## Load Packages 
```{r loadLibs, message=FALSE, warning=FALSE}
library(lubridate)
library(knitr)
library(dplyr)
library(sandwich)
library(MASS)
library(lmtest)
library(quantmod)
library(readxl)
library(moments)
library(sandwich)
library(estimatr)

```

```{r}
library(margins) # For the marginal effects of Logit and Probit
library(randomForest)
library(e1071) # For Naive Bayesian Classifier
library(MASS) # Logit, Probit, LDA, QDA
library(class) # for k-NN
library(dplyr)
library(lmtest)
library(sandwich)
library(tree)
```

## Load Data
```{r message=FALSE,warning=FALSE}
emdata <- read_excel('employment_08_09.xlsx')

head(emdata)
emdata <- na.omit(emdata)
```

# Question A

```{R}
num_employed <- sum(emdata$employed)
frac = num_employed / nrow(emdata)
frac
```
```{R}
se <- sqrt(frac * (1 - frac) / nrow(emdata))

z_score <- qnorm(0.975)

conf_interval_lower <- frac - z_score * se
conf_interval_upper <- frac + z_score * se

conf_interval_lower
conf_interval_upper
```

# Question B

```{R}
emdata$age_squared <- emdata$age^2

binary_lpm <- lm(employed ~ age + age_squared, emdata)
summary(binary_lpm)
coeftest(binary_lpm, vcov = vcovHC(binary_lpm), type = "HC1")
```
```{R}
new_data <- data.frame(age = c(20, 40, 60), age_squared = c(20^2, 40^2, 60^2))

predicted_probs <- predict(binary_lpm, newdata = new_data, type = "response")

predicted_probs
```



```{R}
binary_logit <- glm(employed ~ age + age_squared, family = binomial(link = "logit"), emdata)
binary_probit <- glm(employed ~ age + age_squared, family = binomial(link = "probit"), emdata)

coeftest(binary_logit, vcov = vcovHC(binary_logit), type = "HC1")
coeftest(binary_probit, vcov = vcovHC(binary_probit), type = "HC1")
```


# Question g
```{R}
# Predict probabilities
lm_predictions <- predict(binary_lpm, emdata)

# Convert to classifications
lm_classifications <- ifelse(lm_predictions > 0.5, 1, 0)

# Calculate accuracy
lm_accuracy <- mean(lm_classifications == emdata$employed)
lm_accuracy



# Predict probabilities
probit_predictions <- predict(binary_probit, emdata, type = "response")

# Convert to classifications
probit_classifications <- ifelse(probit_predictions > 0.5, 1, 0)

# Calculate accuracy
probit_accuracy <- mean(probit_classifications == emdata$employed)
probit_accuracy




# Predict probabilities
logit_predictions <- predict(binary_logit, emdata, type = "response")

# Convert to classifications
logit_classifications <- ifelse(logit_predictions > 0.5, 1, 0)

# Calculate accuracy
logit_accuracy <- sum(logit_classifications == emdata$employed) / nrow(emdata)
logit_accuracy
```


# Question h

# Decision Tree
```{R}
# Decision Tree
library(rpart)
decision_tree_model <- rpart(as.factor(employed) ~ age +  I(age ^ 2), data = emdata)

dt_predictions <- predict(decision_tree_model, newdata = emdata, type = "class")

# Accuracy
dt_accuracy <- sum(dt_predictions == emdata$employed) /  nrow(emdata)
dt_accuracy

#BELOW is calculating the precision, recall and F1 score:
dt_predictions_factor <- as.factor(dt_predictions)
employed_factor <- as.factor(emdata$employed)

# Use the confusionMatrix function from the caret package
conf_matrix <- confusionMatrix(dt_predictions_factor, employed_factor)

# Print the confusion matrix along with other statistics
print(conf_matrix)

# Recall (Sensitivity), Precision (Positive Predictive Value), and F1 Score can be extracted as follows:
recall <- conf_matrix$byClass['Sensitivity']
precision <- conf_matrix$byClass['Pos Pred Value']
F1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
print(paste("Recall:", recall))
print(paste("Precision:", precision))
print(paste("F1 Score:", F1_score))
```
From the confusion matrix, we can see that the model just predict every data point as 1, although the accuracy seems high, it is due to the large proportion of 1 in our data. Therefore, we can also see that the recall for class 0 is 0, the precision and F1 score are all NaN.

# Random Forest
```{R}
library(randomForest)
library(caret)

# emdata$employed <- as.character(emdata$employed)
# emdata$employed <- as.factor(emdata$employed)
# random_forest_model <- randomForest(employed ~ ., data = emdata)

set.seed(123) # Setting a seed for reproducibility

rf_model <- randomForest(as.factor(employed) ~ age +  I(age ^ 2), data= emdata, num.trees= 100) 
rf_predictions <- predict(rf_model , emdata)

# Accuracy
rf_accuracy <- mean(rf_predictions == emdata$employed)
rf_accuracy

#BELOW is calculating the precision, recall and F1 score:
rf_predictions_factor <- as.factor(rf_predictions)
employed_factor <- as.factor(emdata$employed)

# Use the confusionMatrix function from the caret package
conf_matrix <- confusionMatrix(rf_predictions_factor, employed_factor)

# Print the confusion matrix along with other statistics
print(conf_matrix)

# Recall (Sensitivity), Precision (Positive Predictive Value), and F1 Score can be extracted as follows:
recall <- conf_matrix$byClass['Sensitivity']
precision <- conf_matrix$byClass['Pos Pred Value']
F1_score <- 2 * (precision * recall) / (precision + recall)

# Print the metrics
print(paste("Recall:", recall))
print(paste("Precision:", precision))
print(paste("F1 Score:", F1_score))
```
Same as above. The model just classify every data point as class 1.


# KNN
```{R}
library(class)

# train_data_scaled <- scale(emdata)
# test_data_scaled <- scale(emdata, center = attr(train_data_scaled, "scaled:center"), scale = attr(train_data_scaled, "scaled:scale"))
# 
# # Fit the k-NN model and make predictions
# k <- 5  # Choose an appropriate k
# knn_pred <- knn(train = train_data_scaled, test = test_data_scaled, cl = train_target, k = k)
# 
# # Evaluate model performance
# table(Predicted = knn_pred, Actual = test_target)

scaled_data <- scale(emdata[, c("age", "age_squared")])
# Define the number of neighbors
k <- 7
knn_predictions <- knn(train = scaled_data, test = scaled_data, cl = as.factor(emdata$employed), k = k)

knn_accuracy    <- sum(knn_predictions == emdata$employed) / nrow(emdata)
knn_accuracy
```

